\documentclass{ximera}
\input{../preamble.tex}

\title{Matrices} \license{CC BY-NC-SA 4.0}
\begin{document}
\begin{abstract}
\end{abstract}
\maketitle

The templates in this section provide sample code for basic matrix operations.  To use Octave, go to the \href{https://sagecell.sagemath.org/}{Sage Math Cell Webpage}, copy the code below into the cell, select OCTAVE as the language, and press EVALUATE.  We also provide a link to our code at the bottom of each template.

\begin{template}\label{temp:matrixOps}

\begin{verbatim}
% Define matrices A, B, and C
A = [1 -1 0 0;
     2 -2 1 2;
     0 1 0 1];
     
B = [2 3 -1 4;
     1 -1 2 -2;
     1 -2 0 3];   
     
C = [1 -2 1 4;
     2 -2 0 1;
     -2 1 1 0;
     1 3 -2 -1];     
     
% Find A+B
sum=A+B

% Find the transpose of A
% Copying and pasting A'  may not work because the prime comes through
% in a different font.
A_trans1=A'
% we can also use the transpose function
A_trans2=transpose(A)

% Find AC
product=A*C

% Find the inverse of C
inv_C=inv(C)

% Find the determinant of C
det_C=det(C)
\end{verbatim}

\href{https://sagecell.sagemath.org/?z=eJxdUEFuwyAQvFvyH-YSNWkTyZDcIg6Yqp-oogjZWEGqIcK4_X4XartOOKyY2ZmdhQ3eTWedQa9jsI0ZIPeo99CuhSoLCYFPhgNDhepcFkiH48DBwGdcIbXZZcZlUScbxzEZTzOdxyTviuBkPF7O2TR51RSZIk4PkZQx49xl_yuxlEUilmetpm3wYekp8q0ui2HsRb4sdLwZxKDdcPeDge8g6cnXzDAhX5Lux6DRDvpr8BhJ9GjpRtdE691i42JpbuVuFSXpN-_Bt2MThXxVT0tY923C3wrUInRVgupW7Z6ErYkm9NZpFycxMSSmSuJfcXxtIA==&lang=octave&interacts=eJyLjgUAARUAuQ==}{Link to code}.
\end{template}

\begin{template}\label{temp:id_matrix}
Sometimes it is useful to generate an identity matrix.  To generate a $5\times 5$ identity matrix, use the code below.
\begin{verbatim}
% n by n identity matrix
I_5=eye(5)
\end{verbatim}
\end{template}

\begin{template}\label{temp:LU}
We can find LU factorization of a matrix as follows.

\begin{verbatim}
    % Define matrix A
A=[2 2 -1 2;
   -1 2 2 1;
   1 2 0 -2;
   1 1 -2 -4];
   
% Find LU factorization for A   
[L, U] = lu (A)

% Is this factorization correct?
L*U
\end{verbatim}

\href{https://sagecell.sagemath.org/?z=eJxdi8EKAiEYhO-C7zCXhYqElG6xByE6ebXLsgcxpR9KwQyqp8_dpUsMDPMNMx2OIVIKuLta6AXNme4HBQUhoQ6cAXNqkgtNsINQP5ItQ-zHhTnrcKJ0gbGIztdc6OMq5YSYC_S8GMwWdkSP2xMrveZsOp1DofhGvbrajB5_71b4XErwlTOzsV_8KS6E&lang=octave&interacts=eJyLjgUAARUAuQ==}{Link to code}.

\begin{remark}
    With the last line of our code we checked the correctness of our factorization.  Did we get the right answer?  Interpret the value $-2.77556e-17$ in the product $LU$.  What do you think caused this value to occur?  
\end{remark}
    
\end{template}

\end{document}