\documentclass{ximera}
\input{../preamble.tex}

\title{Challenge Problems for Ch 9} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\section*{Challenge Problems for Chapter 9}

\begin{problem}\label{prob:nich7.3.10}
    Given linear transformations 
    \begin{center}
\begin{tikzpicture}
\node{
\begin{tikzcd}
\RR^2\rar{T_1}\arrow[black, bend right]{rr}[black,swap]{T}  & \RR^2 \rar{T_2}  & \RR^2
\end{tikzcd}
};
\end{tikzpicture}
\end{center}
\begin{enumerate}
    \item If $T_1$ and $T_2$ are both one-to-one, show that $T$ is one-to-one.
    \item If $T_1$ and $T_2$ are both onto, show that $T$ is onto.
\end{enumerate}
\end{problem}

\begin{problem}\label{prob:7.3.11}
Let $T:V\rightarrow W$ be a linear transformation.
\begin{enumerate}
    \item If $T$ is one-to-one and $TR_1=TR_2$ for linear transformations $R_1,R_2:U\rightarrow V$, show that $R_1=R_2$.
    \item If $T$ is onto and $S_1T=S_2T$ for linear transformations $S_1,S_2:W\rightarrow U$, show that $S_1=S_2$.
\end{enumerate}
    \end{problem}

\begin{problem}\label{prb:10.21} Consider functions defined on $\left\{ 1,2,\cdots ,n\right\} $ having
values in $\mathbb{R}$. Explain how, if $V$ is the set of all such
functions, $V$ can be considered as $\mathbb{R}^{n}$.
\begin{hint}
Let $f\left( i\right) $ be the $i^{th}$ component of a vector $
\vec{x}\in \mathbb{R}^{n}$. Thus a typical element in $\mathbb{R}^{n}$ is $
\left( f\left( 1\right) ,\cdots ,f\left( n\right) \right) $.
\end{hint}
\end{problem}

\begin{problem}\label{prob:nich7.3.13}
    Let $T:v\rightarrow U$ and $S:\rightarrow W$ be linear transformations.
    \begin{enumerate}
        \item If $ST$ is one-to-one, show that $T$ is one-to-one and that $\dim V\leq \dim U$.
        \item If $ST$ is onto, show that $S$ is onto and that $\dim W\leq \dim U$.
    \end{enumerate}
\end{problem}

\begin{problem}\label{prob:nich7.3.18}
    Let $\mathbb{D}_n$ denote the space of all functions $f:\{1, 2, \dots, n\}\rightarrow\RR^n$ (see Problem \ref{prb:10.21}).  If $T:\mathbb{D}_n\rightarrow \RR^n$ is defined by 
    $$T(f)=(f(1), f(2), \dots, f(n))$$
    show that $T$ is an isomorphism.
\end{problem}

\begin{problem}\label{prb:10.6} Let $\mathbb{R}^{\mathbb{N}}$ denote the set of all real valued sequences.
For $\vec{a}\equiv \left\{ a_{n}\right\} _{n=1}^{\infty },\vec{b}\equiv
\left\{ b_{n}\right\} _{n=1}^{\infty }$ two of these, define their sum to be
given by
\begin{equation*}
\vec{a}+\vec{b} =  \left\{ a_{n}+b_{n}\right\} _{n=1}^{\infty }
\end{equation*}
and define scalar multiplication by
\begin{equation*}
c\vec{a}=\left\{ ca_{n}\right\} _{n=1}^{\infty }\text{ where }\vec{a}
=\left\{ a_{n}\right\} _{n=1}^{\infty }
\end{equation*}
Is this a vector space?
\end{problem}

\begin{problem}\label{prb:10.7} Let $\mathbb{C}^{2}$ be the set of ordered pairs of complex numbers.
Define addition and scalar multiplication in the usual way.
\begin{equation*}
\left( z,w\right) +\left( \hat{z},\hat{w}\right) = \left( z+\hat{z},w+
\hat{w}\right) ,\ u\left( z,w\right) \equiv \left( uz,uw\right)
\end{equation*}
Here the scalars are from $\mathbb{C}$. Show this is a vector space.

\end{problem}


\section*{Bibliography}

Several problems come from Chapter 9 of Ken Kuttler's \href{https://open.umn.edu/opentextbooks/textbooks/a-first-course-in-linear-algebra-2017}{\it A First Course in Linear Algebra}. (CC-BY)

Ken Kuttler, {\it  A First Course in Linear Algebra}, Lyryx 2017, Open Edition, pp. 469--535.

Several problems come from the end of Chapter 7 of Keith Nicholson's \href{https://open.umn.edu/opentextbooks/textbooks/linear-algebra-with-applications}{\it Linear Algebra with Applications}. (CC-BY-NC-SA)

W. Keith Nicholson, {\it Linear Algebra with Applications}, Lyryx 2018, Open Edition, pp. 396. 

\end{document}